Sure! Hypothesis testing is a statistical method that helps in making decisions about a population parameter based on a sample statistic. In R, you can perform hypothesis testing using different test statistics such as the z-score, chi-squared statistic, and t-statistic. I'll explain each of these with examples, including the commands for calculating the respective statistic, showing the critical statistic, and comparing them.

### 1. Z-score Test

The z-score test is used when you know the population standard deviation and the sample size is large (n \> 30).

#### Example Scenario:

You want to test if the mean weight of a sample of 50 people is significantly different from the population mean of 70 kg, with a known population standard deviation of 10 kg.

#### Commands:

```{r}
# Sample data
sample_mean <- 72
population_mean <- 70
population_sd <- 10
sample_size <- 50

# Calculate z-score
z_score <- (sample_mean - population_mean) / (population_sd / sqrt(sample_size))
z_score

# Determine critical value for 95% confidence level (two-tailed test)
alpha <- 0.05
z_critical <- qnorm(1 - alpha/2)
z_critical

# Compare z-score with critical value
if (abs(z_score) > z_critical) {
  print("Reject the null hypothesis")
} else {
  print("Fail to reject the null hypothesis")
}
```

### 2. Chi-squared Test

The chi-squared test is used to test the independence of two categorical variables or to test the goodness of fit of a distribution.

#### Example Scenario:

You want to test if the observed frequencies of a categorical variable differ from the expected frequencies.

#### Commands:

```{r}
# Observed and expected frequencies
observed <- c(50, 30, 20)
expected <- c(40, 40, 20)

# Calculate chi-squared statistic
chi_squared_stat <- sum((observed - expected)^2 / expected)
chi_squared_stat

# Determine critical value for 95% confidence level (df = number of categories - 1)
alpha <- 0.05
df <- length(observed) - 1
chi_squared_critical <- qchisq(1 - alpha, df)
chi_squared_critical

# Compare chi-squared statistic with critical value
if (chi_squared_stat > chi_squared_critical) {
  print("Reject the null hypothesis")
} else {
  print("Fail to reject the null hypothesis")
}
```

### 3. T-test

The t-test is used when the population standard deviation is unknown and the sample size is small (n \<= 30).

#### Example Scenario:

You want to test if the mean weight of a sample of 15 people is significantly different from the population mean of 70 kg.

#### Commands:

```{r}
# Sample data
sample <- c(68, 72, 71, 69, 70, 73, 67, 74, 65, 66, 72, 71, 68, 69, 70)
population_mean <- 70

# Perform t-test
t_test <- t.test(sample, mu = population_mean)
t_test

# Extract t-statistic and p-value
t_statistic <- t_test$statistic
p_value <- t_test$p.value

# Determine critical value for 95% confidence level (two-tailed test)
alpha <- 0.05
t_critical <- qt(1 - alpha/2, df = length(sample) - 1)
t_critical

# Compare t-statistic with critical value or p-value with alpha
if (abs(t_statistic) > t_critical) {
  print("Reject the null hypothesis")
} else {
  print("Fail to reject the null hypothesis")
}

# Alternatively, using p-value
if (p_value < alpha) {
  print("Reject the null hypothesis")
} else {
  print("Fail to reject the null hypothesis")
}
```

### Summary of Commands:

-   **Z-score Test:**
    -   `qnorm()`: To find the critical z-value.
    -   `z_score`: Calculation of the z-score.
-   **Chi-squared Test:**
    -   `qchisq()`: To find the critical chi-squared value.
    -   `chi_squared_stat`: Calculation of the chi-squared statistic.
-   **T-test:**
    -   `t.test()`: To perform the t-test.
    -   `qt()`: To find the critical t-value.
    -   `t_test$statistic` and `t_test$p.value`: To extract the t-statistic and p-value from the t-test result.

In all these examples, you compare the calculated statistic with the critical value to decide whether to reject the null hypothesis. Alternatively, for the t-test, you can compare the p-value with the significance level (alpha).
